---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# AICcPermanova

<!-- badges: start -->
[![R-CMD-check](https://github.com/Sustainscapes/AICcPerm/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/Sustainscapes/AICcPerm/actions/workflows/R-CMD-check.yaml)
[![CRAN status](https://www.r-pkg.org/badges/version/AICcPermanova)](https://CRAN.R-project.org/package=AICcPermanova)
[![Codecov test coverage](https://codecov.io/gh/Sustainscapes/AICcPerm/branch/master/graph/badge.svg)](https://app.codecov.io/gh/Sustainscapes/AICcPerm?branch=master)
<!-- badges: end -->

The aim of the AICcPerm repository is to provide an R package that enables model selection for Permanovas from the vegan package. To install this package, you can use the remotes package to install its development version:

```{r, eval= F}
remotes::install_github("Sustainscapes/AICcPerm")
```

Alternatively, you can install the stable version from CRAN:

```{r, eval = F}
install.packages("AICcPermanova")
```

## Generating all possible models

To generate all possible models, you can use the `make_models` function. For example:

```{r examplemakemodels, results='hide', message=FALSE}
library(AICcPermanova)
AllModels <- make_models(vars = c("pH", "Sand", "Clay"), ncores = 2)
```

This function will create a table of all possible models for the specified variables, which you can see in the following table:

```{r Table1, echo = F}
knitr::kable(AllModels)
```

Where you can see all possible models for those 3 variables.

## Calculating AICc

To calculate AICc, you can use the AICc_permanova2 function. Here's an example of how to use it:

```{r}

library(vegan)
data(dune)
data(dune.env)

# Run PERMANOVA using adonis2

Model <- adonis2(dune ~ Management*A1, data = dune.env)

# Calculate AICc
Table_permanova <- AICc_permanova2(Model)
```

The results of this calculation are displayed in the following table:

```{r permanovatable}
knitr::kable(Table_permanova)
```

In this example, we used the adonis2 function to run a PERMANOVA analysis on the dune dataset with the Management and A1 variables. We then calculated AICc using the AICc_permanova2 function

## Full example

In this section, we'll provide a complete example of the AICcPerm package workflow. First, we need to load datasets from the vegan package:

```{r}
library(vegan)
data(dune)
data(dune.env)
```

Next, we'll generate all possible first-order models for this dataset:

```{r, results='hide'}
AllModels <- make_models(vars = c("A1", "Moisture", "Management", "Use", "Manure"))
```

This results in `r nrow(AllModels)` possible models, which are shown in the following table:
```{r, echo=FALSE}
knitr::kable(AllModels)
```

### Avoiding multicollinearity

After generating all the models, it's important to check for multicollinearity. We can use the `filter_vif` function to filter out models that have a high degree of collinearity (defined as having a maximum value of VIF of 5 or more):

```{r}
NonColinear <- filter_vif(all_forms = AllModels, env_data = dune.env)
```

This reduces the number of models to `r nrow(NonColinear)`

### Fittng the models

After filtering out collinear models, we can fit all the remaining non-collinear models by using the `fit_models` function:

```{r}
Fitted <- fit_models(all_forms = NonColinear,
           veg_data = dune,
           env_data = dune.env,
           ncores = 4,
           method = "bray")
```

This results in a table of fitted models ordered by AICc, which is displayed below:

```{r, echo=FALSE}
knitr::kable(Fitted)
```

If there is a block variable to be used (such as Use in the dune.env object), you can specify it using the strata argument:

```{r fitwihtstrata}
Fitted2 <- fit_models(all_forms = NonColinear,
           veg_data = dune,
           env_data = dune.env,
           ncores = 4,
           method = "bray",
           strata = "Use")
```

This results in a table of fitted models that takes into account the block variable, which is displayed below:

```{r, echo=FALSE}
knitr::kable(Fitted2)
```

## Model Selection

To select models, we use the select_models function, which chooses models with a maximum VIF of less than 5 and a delta AICc less than a specified threshold, delta_aicc. By default, delta_aicc is set to 2.

We can use the Fitted2 object generated in the previous section to select models:

```{r selected}
Selected <- select_models(Fitted2)
```

The resulting table displays the selected models:

```{r tableselected, echo = FALSE}
knitr::kable(Selected)
```

Note that the models in the table satisfy the criteria for maximum VIF and delta AICc as specified in the `select_models` function.


### Summary weighted by AICc

finally you can do a summarized r squared weighted by AICc using the 
`akaike_adjusted_rsq` function as seen bellow:

```{r}
Summary <- akaike_adjusted_rsq(Selected)
```

which results in the following table:

```{r, echo = FALSE}
knitr::kable(Summary)
```

